Donor Preferences in Canadian Political Financing

### 2024-12-15 paper/paper.qmd
renamed folder
update model to use real_amount

# Aggregate total donations per party per year in Ontario
total_donations_ontario <- ontario_data %>%
  group_by(political_party, donation_year) %>%
  summarize(
    Total_Donations = sum(amount, na.rm = TRUE),
    In_Power = first(recipient_in_power),
    Party_Size = first(Party_Size),
    Election_Year = first(Election_Year)
  ) %>%
  ungroup()

@@

# Aggregate total donations per party per year at Federal level
total_donations_federal <- federal_data %>%
  group_by(political_party, donation_year) %>%
  summarize(
    Total_Donations = sum(amount, na.rm = TRUE),
    In_Power = first(recipient_in_power),
    Party_Size = first(Party_Size),
    Election_Year = first(Election_Year)
  ) %>%
  ungroup()

@@

# Aggregate total donations per party per year
total_donations_cl <- cl_data %>%
  group_by(political_party, donation_year) %>%
  summarize(
    Total_Donations = sum(amount, na.rm = TRUE),
    In_Power = first(In_Power),
    Party = first(Party),
    Election_Year = first(Election_Year)
  ) %>%
  ungroup()

@@

In the Ontario model, Party_Size is a significant predictor of total
donations, with a coefficient of 2.956 (standard error = 0.485, p \<
0.001). This indicates that major parties in Ontario receive
significantly higher total donations than minor parties, highlighting
the influence of party prominence on fundraising success.

The In_Power variable has a coefficient of 0.884 (standard error =
0.540, p = 0.106), which is not statistically significant at the
conventional 5% level. This suggests that being the ruling party in
Ontario does not have a significant direct effect on the total amount of
donations received.

The Election_Year variable is also not significant in the Ontario model
(coefficient = 0.239, standard error = 0.436, p = 0.586), indicating
that donations are not substantially influenced by whether they occur
during an election year.

@@

In the Federal model, Party_Size remains a significant predictor, with a
coefficient of 3.450 (standard error = 0.259, p \< 0.001). This
reinforces the importance of party size in attracting donations at the
federal level.

The In_Power variable in the Federal model has a coefficient of 0.637
(standard error = 0.3882, p = 0.103), which is not statistically
significant. Similar to the Ontario context, being in power at the
federal level does not have a significant direct effect on the total
donations received.

The Election_Year variable is not significant in the Federal model
(coefficient = -0.0752, standard error = 0.2209, p = 0.734), suggesting
no substantial impact of election years on donation amounts at the
federal level.

@@

In the Conservative/Liberal model, In_Power has a significant positive
effect, with a coefficient of 1.202 (standard error = 0.497, p =
0.01827). Additionally, the interaction term In_Power × Party is
significantly negative, with a coefficient of -1.718 (standard error =
0.711, p = 0.01845). This interaction indicates that the effect of being
in power on total donations varies between the Conservative and Liberal
parties.

For the Conservative Party (Party = 1), the combined effect of being in
power is 1.2023 + (-1.7175) = -0.5152, suggesting that being in power is
associated with lower total donations for the Conservative Party. For
the Liberal Party (Party = 0), the effect of being in power remains
1.2023, indicating that being in power is associated with higher total
donations for the Liberal Party.

The Party variable itself is significant (coefficient = 1.3648, standard
error = 0.4968, p = 0.00774), suggesting that, overall, the Conservative
Party tends to receive more donations than the Liberal Party when not
considering the interaction effect.

The Election_Year variable is not significant in the
Conservative/Liberal model (coefficient = -0.0527, standard error =
0.3777, p = 0.88946).

@@

@fig-estimates highlights the consistently strong positive effect of
Party_Size on total donations in both the Ontario and Federal models,
emphasizing that major parties attract significantly more donations than
minor parties. It also illustrates the significant interaction effect
between In_Power and Party in the Conservative/Liberal model,
demonstrating the differing impact of being in power on donations for
the two major parties.

In summary, the analyses reveal that party size is a significant
determinant of total donations at both provincial and federal levels,
while the effect of being in power on donations is contingent upon party
affiliation, particularly between the Conservative and Liberal parties.

>>>>

# Aggregate total donations per party per year in Ontario
total_donations_ontario <- ontario_data %>%
  group_by(political_party, donation_year) %>%
  summarize(
    Total_Donations = sum(real_amount, na.rm = TRUE),
    In_Power = first(recipient_in_power),
    Party_Size = first(Party_Size),
    Election_Year = first(Election_Year)
  ) %>%
  ungroup()

@@

# Aggregate total donations per party per year at Federal level
total_donations_federal <- federal_data %>%
  group_by(political_party, donation_year) %>%
  summarize(
    Total_Donations = sum(real_amount, na.rm = TRUE),
    In_Power = first(recipient_in_power),
    Party_Size = first(Party_Size),
    Election_Year = first(Election_Year)
  ) %>%
  ungroup()

@@

# Aggregate total donations per party per year
total_donations_cl <- cl_data %>%
  group_by(political_party, donation_year) %>%
  summarize(
    Total_Donations = sum(real_amount, na.rm = TRUE),
    In_Power = first(In_Power),
    Party = first(Party),
    Election_Year = first(Election_Year)
  ) %>%
  ungroup()

@@

In the Ontario model, Party_Size is a significant predictor of total
donations, with a coefficient of 3.046 (standard error = 0.499, p \<
0.001). This indicates that major parties in Ontario receive
significantly higher total donations than minor parties, highlighting
the influence of party prominence on fundraising success.

The In_Power variable has a coefficient of 0.886 (standard error =
0.547, p = 0.116), which is not statistically significant at the
conventional 5% level. This suggests that being the ruling party in
Ontario does not have a significant direct effect on the total amount of
donations received.

The Election_Year variable is also not significant in the Ontario model
(coefficient = 0.274, standard error = 0.449, p = 0.544), indicating
that donations are not substantially influenced by whether they occur
during an election year.

@@

At the federal level, Party_Size once again emerges as a significant determinant of total donations, with a coefficient of 3.452 (standard error = 0.259, p \< 0.001). This confirms that larger, more established parties are more successful at fundraising on the national stage, similar to the provincial setting.

The In_Power variable is not statistically significant in the federal model (coefficient = 0.637, standard error = 0.389, p = 0.103), implying that being the governing party at the federal level does not confer a clear advantage in terms of donation totals.

Similarly, Election_Year does not significantly impact federal donations (coefficient = -0.042, standard error = 0.221, p = 0.849), indicating no substantial change in donor behavior during election years at the federal level.

@@

In the Conservative/Liberal model, the In_Power variable is now positively and significantly related to total donations (coefficient = 1.201, standard error = 0.513, p = 0.022), while the interaction In_Power × Party is significantly negative (coefficient = -1.711, standard error = 0.734, p = 0.023). This interaction means that the effect of being in power depends on whether the party is Conservative or Liberal.

For the Conservative Party (Party = 1), the combined effect of being in power is approximately 1.201 + (-1.711) = -0.510, suggesting that being in power is associated with lower total donations for the Conservative Party. For the Liberal Party (Party = 0), the effect of being in power remains 1.201, indicating that being in power is linked to higher donation totals for the Liberal Party.

The Party variable itself is also significant (coefficient = 1.364, standard error = 0.513, p = 0.0099), indicating that, when not considering the interaction, the Conservative Party tends to attract more donations than the Liberal Party. The Election_Year variable remains non-significant (coefficient = -0.023, standard error = 0.390, p = 0.953), suggesting no direct election-year influence on donations within this subset.

@@

@fig-estimates continues to highlight the consistently strong positive effect of Party_Size on total donations in both the Ontario and Federal models, affirming that well-established parties attract more donor support. It also illustrates the significant interaction effect between In_Power and Party in the Conservative/Liberal model, demonstrating that the influence of being in power on donations differs notably between the two major parties.

In summary, the updated results show that party size consistently drives higher donations across jurisdictions, while the impact of being in power on donation amounts depends on the party’s ideological affiliation, especially when contrasting Conservative and Liberal parties.
